Seeing Double: Transient Dips and Photometric Binarity in PTFO$\,$8-8695

### 2020-04-23 paper/ms.tex
better conclusion

\documentclass[12pt,twocolumn,tighten]{aastex62}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat

@@

\begin{abstract}
  PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10 million year
  old Orion-OB1a cluster. We inspected data from TESS and Gaia to
  clarify whether it is truly a planet.  The Gaia data show that
  PTFO$\,$8-8695 is a photometric binary with respect to members of
  its kinematic group.  The TESS lightcurve shows that the dominant
  variability in this system is a sinusoid with a long period
  $P_{\rm \ell}=11.96\,$hr, presumably caused by stellar rotation.
  Also present is a complex signal, previously identified as the
  planet candidate, that repeats with a short period $P_{\rm s}=
  10.74\,$hr.  The two signals beat every 4.48 days.  Although there
  is a dip in the short-period signal, ground-based photometry from
  the past decade indicates that the orbital phase of the dip seems to
  have instantaneously jumped, at least once, and perhaps twice.
  Planets do not ``jump'' in orbital phase.  Given the available
  evidence, we believe that PTFO$\,$8-8695 is a binary M dwarf in
  which one star shows the long rotation signal, and the other star
  shows ``transient dipping'' that has been seen in a few other young
  weak-lined T Tauri stars.  The origin of these transient dips is
  still undetermined, but our preferred explanation is eclipsing
  clouds of gas or dust at the Keplerian co-rotation radius.
\end{abstract}

@@

We begin in Section~\ref{sec:observations} by describing newly
available observations from TESS \citep{ricker_transiting_2015} and
Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows
two distinct signals, which we extract and analyze in
Section~\ref{sec:tess}.  A long-period sinusoid repeats every 12.0
hours, and is probably stellar rotation.  A short-period dip with an
additional complex modulation repeats every 10.7 hours.  Analyzing the
Gaia data in Section~\ref{sec:gaia}, we show that relative to its
kinematic group, \ptfo\ is a photometric binary.  We collect and
discuss the puzzle pieces in Section~\ref{sec:discussion}.  The
orbital phase of the dip seems to have instantaneously changed over
the past decade.  In addition, a number of other young stars show
lightcurve morphologies similar to the short-period signal.  We
therefore argue that PTFO$\,$8-8695 is a binary M dwarf in which one
star shows a rotation signal, and the other shows ``transient
dipping'' caused by eclipses of material at the Keplerian co-rotation
radius.  Section~\ref{sec:conclusions} summarizes our main points.

@@

The majority of Gaia's derived parameters for \ptfo\ agreed with
expectation from former studies
\citep{briceno_cida_2005,van_eyken_ptf_2012}.  One novelty however was
that Gaia DR2 detected a significant ``astrometric excess'', at a
level of 10.3$\sigma$.  We comment on the significance and
interpretation of this excess in Section~\ref{sec:gaia}.

@@

We implemented and fitted the models using \texttt{PyMC3}, which is
built on \texttt{theano}
\citep{salvatier_2016_PyMC3,exoplanet:theano}.  For the Fourier terms,
we used the default math operators.  For the exoplanet transit, we
used the model and derivatives implemented in \texttt{exoplanet}
\citep{exoplanet:exoplanet}.  Our priors are listed in Table~2.  To
speed up the fitting, we re-sampled the cleaned 2 minute lightcurves
to 10 minute binning.  We correspondingly scaled the uncertainties in
the flux measurements by a factor of $\sqrt{5}$.  Before sampling, we
initialized each model to the maximum a posteriori (MAP) solution.  We
then sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
\citep{hoffman_no-u-turn_2014}, and used $\hat{R}$ as our convergence
diagnostic \citep{gelman_inference_1992}.  We tested our ability to
successfully recover injected parameters using synthetic data, before
fitting the \ptfo\ lightcurves.

@@

In every model, the $11.96\,{\rm hr}$ variability is a simple sinusoid
with peak-to-peak amplitude $\approx$10\%.  The $10.74\,{\rm hr}$
variability is always more complex.  A dip of depth $\approx$1.2\%,
fit in our model as a transit, lasts $\approx$0.75 hours.  Superposed
on the dip is a slightly asymmetric sinusoid with peak-to-peak
amplitude of about 4\%. The asymmetric sinusoid peaks near phase 0.25,
and reaches minimum brightness between phases -0.5 and -0.25.  The
flux at phase $\approx -0.33$ shows what could be a discontinuous
jump, shortly after reaching minimum (Figure~\ref{fig:phasefold}).
This jump was visible in each of the nine models we considered.

The periodogram of the final residual (Figure~\ref{fig:splitsignal}
bottom row) shows a weakly significant, poorly resolved peak at
$\approx$8 days, consistent with the visual impression in the time
domain that there could be a weak long-period signal present.

@@

% NB: CDIPS-I only made 131 light-curves for 25Ori-1, because I wasn't
% aware of the Kounkel+18 member list at the time.
Using the same 149 members in the 25$\,$Ori-1 subgroup, we calculated
the astrometric reduced $\chi^2$ for each member.  We then queried the
CDIPS lightcurve database at MAST \citep{bouma_cluster_2019} to find
the subset of stars that were at least as variable as \ptfo.
We measured the variability amplitude by taking the difference of the
$95^{\rm th}$ and $5^{\rm th}$ percentiles of the flux.
This yielded 30 stars of equal or greater variability.
The lower panel of Figure~\ref{fig:gaia} shows the
reduced $\chi^2$ as a function of stellar brightness.  \ptfo\ is in
the upper 90$^{\rm th}$ percentile of stars showing astrometric
excesses within the 25$\,$Ori-1 group.  Relative to other M-dwarf
group members with comparable brightnesses and variability
characteristics, \ptfo\ still stands out as behaving astrometrically
poorly.  Ultimately, we will have to wait for the full release of the
nominal Gaia mission to definitively determine whether the astrometric
excess is caused by stellar binarity or photometric variability.
Nonetheless the fact that comparably variable stars do not show
comparably large astrometric excesses suggests that stellar binarity
is indeed the root cause.

@@

\subsubsection{Similar Lightcurves}

@@

These lightcurves have been phenomenologically classified as
``persistent flux dips'' or ``transient flux dips'', based on whether
their depths and durations show variability over the 90-day K2
campaigns \citep{stauffer_orbiting_2017}.
As \citet{stauffer_orbiting_2017} explain,
these  objects are morphologically distinct from ``scallop shell' lightcurves,
and are present in stars at more advanced evolutionary disk stages than
the ``dipper'' stars \citep{ansdell_young_2016,cody_manyfaceted_2018}.
The persistent and transient flux dip stars have the following properties.
\begin{enumerate}
  \item Weak-lined T Tauri stars.
  \item Spectral type typically M2-M5 ({\it e.g.},
    \citealt{rebull_usco_2018},~Figure~20).
  \item Ages $\lesssim$ 100$\,$Myr.
  \item Stars typically lack spectroscopic accretion indicators 
  \item Phased lightcurves show shallow, angular dips, usually
    superposed on large-amplitude smooth variability. The latter
    is interpreted as stellar rotation.
  \item Rotation is fast---usually between 0.5 and 2.0 days.
  \item Rarely any infrared excess (never any W4 detection; only a few
    cases with W3 detctions).
  \item Can show multiple dips per cycle ({\it e.g.}, EPIC 203692610,
    RIK-210).
  \item Dip depths, durations, and phases can vary over just a few cycles
    ({\it e.g.}, EPIC 204143627).
  \item Dip depths can change after flares.
  \item Rare at a population level: $\lesssim 1\%$ of relevant stars
    \citep{rebull_usco_2018}.
\end{enumerate}
The 10.74 hour signal in \ptfo\ meets all of these criteria.
This is the first connection of \ptfo\ with this class of objects
because the TESS lightcurve was needed to resolve the different
rotation signals.

@@

\section{Physical Interpretation}

The evidence for binarity in \ptfo\ is as follows.  First, the star is
photometrically twice as bright as stars of the same color in its
kinematic group (Figure~\ref{fig:gaia}.  Second, it shows two distinct
photometric signals.  These points alone suggest binarity
\citep{stauffer_rotevol_2018}.  For the case of \ptfo, there is a
third line of evidence: the Gaia astrometry shows a poor
goodness-of-fit to a single-source model.  While this could be caused
by an additional companion or stellar variability, cluster members
that are equally if not more variable still show \ptfo\ to be an
outlier.  Therefore the astrometric excess is a suggestive third line
of evidence for binarity in \ptfo.  We take these three points
combined to imply that \ptfo\ is probably a binary.

@@

The main physical question remaining is clearly what causes the
transient dipping. This is an unsolved problem not only for \ptfo\ but
also for an emerging class of similar young rapidly rotating M-dwarfs.
Some possible explanations are highly disfavored, for reasons that
have been detailed by \citet{david_transient_2017},
\citet{stauffer_orbiting_2017}, and \citet{zhan_complex_2019}.
These disfavored explanations include that the dips are caused by
{\it (i)} eclipsing binaries;
{\it (ii)} ``dipper''-flavor Class-I or Class-II disks;
{\it (iii)} eclipses of prominences;
{\it (iv)} high-latitude accretion hotspots;
{\it (v)} high-latitude starspots;
and
{\it (vi)} dust clouds of reasonable composition.
We also view the possibility of {\it (vii)} tidally disrupted planetary
or cometary material to be implausible, given the synchronicity
between dip and rotation periods, and the stability of signals
observed in some systems.

The explanations that are not yet ruled out include
{\it (i)} transiting clumps of gas at the Keplerian corotation radius;
{\it (ii)} transits of enshrouded protoplanets;
{\it (iii)} occultations of starspots by an optically thick disk.
The first and last explanations offer added appeal because they are
flexible enough to explain not only the transient and persistent-dip
M-dwarfs, but also the ``scallop shell'' M-dwarfs
\citep{stauffer_orbiting_2017}.
Despite this appeal, the possibility of distinct mechanisms 
explaining these
distinct variability classes should remain open.

@@

\ptfo\ was previously thought to potentially host a hot Jupiter.
The TESS lightcurve of \ptfo\ showed a number of new features,
many of which seem to disfavor the hot Jupiter interpretatation.
The TESS data showed two key pieces of evidence.
\begin{enumerate}
  \item {\it Two periodic signals.} The ``long'' signal is a 10\%
      peak-to-peak sinusoidal modulation repeating every 11.96 hours.
      The ``short'' signal is a 4\% peak-to-peak complex modulation
      repeating every 10.74 hours. It is composed of a dip, plus at
      least two harmonics. The signals beat, and therefore cannot be
      an artifact linked to data processing.
  \item {\it A dip at the wrong orbital phase.} The clearest dip in
    the ``short'' signal was consistent with recent observations by
    \citet{tanimoto_evidence_2020}, and differed from the discovery
    epoch by 5.14 hours.
\end{enumerate}

The physical mechanism responsible for all these features remains a
matter of speculation.
With that said, the TESS data support new arguments against the planetary
interpretation of \ptfo.
First, if the long signal is caused by starspot modulation, and the
short signal by a transiting planet, what causes the additional
complex modulations seen at the short, ``orbital'', period?

Given the available evidence,
PTFO$\,$8-8695 seems consistent with the
``transient dipping'' phenomenology observed in many young M dwarfs.
Though it seems rather unlikely to be a planet, understanding \ptfo\
and its analogs will help us understand the natal environments 
of the majority of Earth-sized planets in the Milky Way
\citep{dressing_occurrence_2013}.

@@

% \acknowledgements
% %
% This paper includes data collected by the TESS mission, which are
% publicly available from the Mikulski Archive for Space Telescopes
% (MAST).
% %
% Funding for the TESS mission is provided by NASA's Science Mission
% directorate.
% %
% This work made use of NASA's Astrophysics Data System Bibliographic
% Services.
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  % \texttt{astroplan} \citep{astroplan2018},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  % \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:agol19}
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol19, exoplanet:kipping13, exoplanet:luger18,
  	exoplanet:theano}.
  \texttt{IPython} \citep{perez_2007},
	\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  \texttt{radvel} \citep{fulton_radvel_2018},
  % \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
}

% \facilities{
% 	{\it Astrometry}:
% 	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}.
% 	{\it Imaging}:
% 	Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
% 	{\it Spectroscopy}:
% 	Keck:I~(HIRES; \citealt{vogt_hires_1994}),
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
% 	{\it Photometry}:
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
% 	El Sauce:0.356m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
% 	NTT,
% 	SOAR (SOI),
% 	TESS \citep{ricker_transiting_2015},
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
% }

@@

\clearpage
\input{model_comparison_table.tex}
\input{model_posterior_table.tex}
\clearpage

>>>>

\documentclass[12pt,twocolumn,tighten]{aastex62}
%\documentclass[12pt,twocolumn,tighten,trackchanges]{aastex62}
\usepackage{amsmath,amstext,amssymb}
\usepackage[T1]{fontenc}
\usepackage{apjfonts}
\usepackage[figure,figure*]{hypcap}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage[caption=false]{subfig} % for subfloat
\usepackage{enumitem} % for specific spacing of enumerate

@@

\begin{abstract}
  PTFO$\,$8-8695b is a candidate hot Jupiter in the 7--10 million year
  old Orion-OB1a cluster. We inspected data from TESS and Gaia to
  clarify whether it is truly a planet.  The Gaia data show that
  PTFO$\,$8-8695 is a photometric binary with respect to members of
  its kinematic group, and also suggest that it is an astrometric
  binary.  The TESS lightcurve shows that the dominant variability in
  this system is a sinusoid with a long period $P_{\rm
  \ell}=11.96\,$hr, presumably caused by stellar rotation.  Also
  present is a combined dip and asymmetric sinusoid, previously
  identified as the planet candidate, that repeats with a short period
  $P_{\rm s}= 10.74\,$hr.  The two signals beat every 4.48 days.
  Although there is a dip in the short-period signal, ground-based
  photometry from the past decade indicates that the orbital phase of
  the dip seems to have instantaneously jumped, at least once, and
  perhaps twice.  Planets do not ``jump'' in orbital phase.  Given the
  available evidence, we believe that PTFO$\,$8-8695 is a binary M
  dwarf in which one star shows the long rotation signal, and the
  other star shows ``transient dipping'' that has been seen in a few
  other young low-mass stars.  The origin of these transient dips is
  still undetermined.
\end{abstract}

@@

We begin in Section~\ref{sec:observations} by describing newly
available observations from TESS \citep{ricker_transiting_2015} and
Gaia \citep{gaia_collaboration_gaia_2018}.  The TESS lightcurve shows
two distinct signals, which we extract and analyze in
Section~\ref{sec:tess}.  A long-period sinusoid repeats every 12.0
hours, and is probably stellar rotation.  A short-period dip superposed on an
asymmetric sinusoid repeats every 10.7 hours.  Analyzing the
Gaia data in Section~\ref{sec:gaia}, we show that relative to its
kinematic group, \ptfo\ is a photometric binary, and suggest it also an
astrometric binary.  We collect and discuss the puzzle pieces in
Section~\ref{sec:discussion}.  The orbital phase of the dip seems to
have instantaneously changed over the past decade.  In addition, a
number of other young stars show lightcurve morphologies similar to
the short-period signal.  We therefore argue that PTFO$\,$8-8695 is a
binary M dwarf in which one star shows a rotation signal, and the
other shows ``transient dipping'' potentially caused by eclipses of plasma at
the Keplerian co-rotation radius.  Section~\ref{sec:conclusions}
summarizes our main points.

@@

The majority of Gaia's derived parameters for \ptfo\ agreed with
expectations from former studies
\citep{briceno_cida_2005,van_eyken_ptf_2012}.  One novelty however was
that Gaia DR2 detected a 10.3$\sigma$ ``astrometric excess''.
We comment on the significance and
interpretation of this excess in Section~\ref{sec:gaia}.

@@

We implemented and fitted the models using \texttt{PyMC3}, which is
built on \texttt{theano}
\citep{salvatier_2016_PyMC3,exoplanet:theano}.  For the Fourier terms,
we used the default math operators.  For the exoplanet transit, we
used the model and derivatives implemented in \texttt{exoplanet}
\citep{exoplanet:exoplanet}.  Our priors are listed in Table~2.  To
speed up the fitting, we re-sampled the cleaned 2 minute lightcurves
to 10 minute binning.  We correspondingly scaled the uncertainties in
the flux measurements by a factor of $\sqrt{5}$.  Before sampling, we
initialized each model to the maximum a posteriori (MAP) solution.  We
then sampled using \texttt{PyMC3}'s gradient-based No-U-Turn Sampler
\citep{hoffman_no-u-turn_2014}, and used $\hat{R}$ as our convergence
diagnostic \citep{gelman_inference_1992}.  We tested our ability to
successfully recover injected parameters using synthetic data before
fitting the \ptfo\ lightcurves.

@@

In every model, the $11.96\,{\rm hr}$ variability is a simple sinusoid
with peak-to-peak amplitude $\approx$10\%.  The $10.74\,{\rm hr}$
variability is always more complex.  A dip of depth $\approx$1.2\%,
fit in our model as a transit, lasts $\approx$0.75 hours.  Superposed
on the dip is an asymmetric sinusoid with peak-to-peak
amplitude of about 4\%. The asymmetric sinusoid peaks near phase 0.25,
and reaches minimum brightness between phases -0.5 and -0.25.  The
flux at phase $\approx -0.33$ shows what could be a discontinuous
jump, shortly after reaching minimum (Figure~\ref{fig:phasefold}).
This jump was visible in each of the nine models we considered.

The periodogram of the final residual (Figures~\ref{fig:splitsignal} and~\ref{fig:splitsignalii}
bottom row) shows a weakly significant, poorly resolved peak at
$\approx$8 days, consistent with the visual impression in the time
domain that there could be a weak long-period signal present.

@@

% NB: CDIPS-I only made 131 light-curves for 25Ori-1, because I wasn't
% aware of the Kounkel+18 member list at the time.
Using the same 149 members in the 25$\,$Ori-1 subgroup, we calculated
the astrometric reduced $\chi^2$ for each member.  We then queried the
CDIPS lightcurve database at MAST \citep{bouma_cluster_2019} to find
the subset of members that were at least as variable as \ptfo.
We measured the variability amplitude by taking the difference of the
$95^{\rm th}$ and $5^{\rm th}$ percentiles of the flux.
This yielded 30 stars of equal or greater variability.
The lower panel of Figure~\ref{fig:gaia} shows the
reduced $\chi^2$ as a function of stellar brightness.  \ptfo\ is in
the upper 90$^{\rm th}$ percentile of stars showing astrometric
excesses within the 25$\,$Ori-1 group.  Relative to other M-dwarf
group members with comparable brightnesses and variability
characteristics, \ptfo\ still stands out as behaving astrometrically
poorly.  Ultimately, we will have to wait for the full release of the
nominal Gaia mission to definitively determine whether the astrometric
excess is caused by stellar binarity or photometric variability.
Nonetheless the fact that comparably variable stars do not show
comparably large astrometric excesses suggests that stellar binarity
is indeed the root cause.

@@

\subsubsection{Similar Lightcurves}
\label{subsec:dipstars}

@@

These lightcurves have been phenomenologically classified as
``persistent flux dips'' or ``transient flux dips'', based on whether
their depths and durations show variability over the 90-day K2
campaigns \citep{stauffer_orbiting_2017}.
As \citet{stauffer_orbiting_2017} explain,
these  objects are morphologically distinct from ``scallop shell' lightcurves,
and are present in stars at more advanced evolutionary disk stages than
the ``dipper'' stars \citep{ansdell_young_2016,cody_manyfaceted_2018}.
The persistent and transient flux dip stars have the following properties.
\begin{enumerate}[topsep=0.5ex,itemsep=-0.5ex,partopsep=1ex,parsep=1.5ex]
  \item Weak-lined T Tauri stars.
  \item Spectral type typically M2-M5 ({\it e.g.},
    \citealt{rebull_usco_2018},~Figure~20).
  \item Ages $\lesssim$ 100$\,$Myr.
  \item Stars typically lack spectroscopic accretion indicators 
  \item Phased lightcurves show shallow, angular dips, usually
    superposed on large-amplitude smooth variability. The latter
    is interpreted as stellar rotation.
  \item Rotation is fast---usually between 0.5 and 2.0 days.
  \item Rarely any infrared excess (never any W4 detection; only a few
    cases with W3 detctions).
  \item Can show multiple dips per cycle.
  %({\it e.g.}, EPIC 203692610, RIK-210).
  \item Dip depths, durations, and phases can vary over just a few cycles
    ({\it e.g.}, EPIC 204143627).
  \item Dip depths can change after flares.
  \item Rare at a population level: $\lesssim 1\%$ of relevant stars
    \citep{rebull_usco_2018}.
\end{enumerate}
The 10.74 hour signal in \ptfo\ meets all of these criteria.
This is the first connection of \ptfo\ with this class of objects
because the TESS lightcurve was needed to resolve the different
rotation signals.

@@

\subsection{Physical Interpretation}
\label{subsec:physical}

The evidence for binarity in \ptfo\ is as follows.  First, the star is
photometrically twice as bright as stars of the same color in its
kinematic group (Figure~\ref{fig:gaia}).  Second, it shows two distinct
photometric signals.  These points alone suggest binarity
\citep{stauffer_rotevol_2018}.  For the case of \ptfo, there is a
third line of evidence: the Gaia astrometry shows a poor
goodness-of-fit to a single-source model.  While this could be caused
by stellar variability, cluster members
that are at least as variable still show \ptfo\ to be an
outlier.  Therefore the astrometric excess is a suggestive third line
of evidence for binarity in \ptfo.  We take these three points
combined to imply that \ptfo\ is probably a binary.

@@

The main physical question remaining is clearly what causes the
transient dipping. This is an unsolved problem not only for \ptfo\ but
also for an emerging class of similar young rapidly rotating M-dwarfs.
Many possible explanations been detailed by \citet{rebull_rotation_2016}, \citet{david_transient_2017},
\citet{stauffer_orbiting_2017}, and \citet{zhan_complex_2019}.
Their disfavored explanations include that the dips are caused by
{\it (i)} eclipsing binaries;
{\it (ii)} ``dipper''-flavor Class-I or Class-II disks;
{\it (iii)} eclipses of prominences;
{\it (iv)} high-latitude accretion hotspots;
{\it (v)} high-latitude starspots;
or
{\it (vi)} dust clouds of plausible composition.
We also view the possibility of {\it (vii)} tidally disrupted planetary
or cometary material to be implausible, given the synchronicity
between dip and rotation periods seen across many systems.

The explanations that are not yet ruled out include
{\it (i)} transiting clumps of gas at the Keplerian corotation radius;
{\it (ii)} transits of enshrouded protoplanets;
{\it (iii)} occultations of starspots by an optically thick disk.
The first and last explanations offer added appeal because they are
flexible enough to explain not only the transient and persistent-dip
M-dwarfs, but also the ``scallop shell'' M-dwarfs
\citep{stauffer_orbiting_2017}.
Despite this appeal, the possibility of distinct mechanisms explaining
these distinct variability classes remains open.

The evolution of \ptfo\
over the past decade (Figure~\ref{fig:o_minus_c}) could offer important
hints.  Specifically, \ptfo's transition between having none, one, and
multiple dips per cycle seems important.  It strains the ``enshrouded
protoplanet'' interpretation, because there are no known processes
that cause a planet's orbital phase to jump.  The dips would then need
to be caused by material that was somehow disrupted from the planet,
but somehow remained co-orbital for an extended duration. This seems
implausible.

@@

The combination of TESS and Gaia observations have provided new
clarity in the \ptfo\ system.  Our main results are as follows.
\begin{itemize}
  \item {\it The TESS light-curve shows two periodic signals.} The
    ``long'' signal is a 10\% peak-to-peak sinusoid that repeats every
    11.96 hours.  The ``short'' signal is a 4\% peak-to-peak ``dip +
    asymmetric sinusoid'' that repeats every 10.74 hours. The signals
    beat, and therefore cannot be an artifact linked to data
    processing.
    Within the angular resolution of the Gaia source
    catalog, both signals need to originate from \ptfo.
  \item {\it The Gaia data imply binarity.} Relative to stars in its
    kinematic group, \ptfo\ is a photometric binary
    (Figure~\ref{fig:gaia}, top).  Relative to stars in its group that
    are at least as photometrically variable, \ptfo\ also shows
    signs of astrometric binarity (Figure~\ref{fig:gaia}, bottom).
  \item {\it The orbital phase of the dip has changed since the
    discovery by \citet{van_eyken_ptf_2012}.} As shown in
    Figure~\ref{fig:o_minus_c}, the phase seems to have jumped,
    perhaps twice. This agrees with the recent study by
    \citet{tanimoto_evidence_2020}.
  \item {\it All properties of \ptfo\ are consistent with the emerging
    class of transient and persistent flux dip stars.} Analogous
    lightcurves are shown in Figure~\ref{fig:brethren}.  Properties of
    this variability class are enumerated in
    Section~\ref{subsec:dipstars}.
\end{itemize}

The physical mechanism that explains the transient and persistent flux
dips is unresolved. Our preferred explanations include
transiting clumps of gas at the Keplerian corotation radius, and
occultations of starspots by an optically thick disk \citep[{\it
e.g.},][]{stauffer_orbiting_2017,david_transient_2017,zhan_complex_2019}.
The jumping orbital phase disfavors the enticing explanation of an
enshrouded, transiting protoplanet.  Though \ptfob seems unlikely
to be a planet, understanding it and its analogs is a worthy
problem that promises to teach us about the natal environments of the
majority of habitable-zone Earth-sized planets in the Milky Way
\citep{dressing_occurrence_2013}.

@@

\acknowledgements
%
%This paper includes data collected by the TESS mission, which are
%publicly available from the Mikulski Archive for Space Telescopes
%(MAST).
%
%Funding for the TESS mission is provided by NASA's Science Mission
%directorate.
%
\ptfo\ was included on the TESS ``short-cadence'' target list thanks to the
Guest Investigator programs of S.\ Czesla and C.\ Huang (G011128 and
G011132 respectively).
%
% %
% Based on observations obtained at the Gemini Observatory, which is
% operated by the Association of Universities for Research in Astronomy,
% Inc., under a cooperative agreement with the NSF on behalf of the
% Gemini partnership: the National Science Foundation (United States),
% National Research Council (Canada), CONICYT (Chile), Ministerio de
% Ciencia, Tecnolog\'{i}a e Innovaci\'{o}n Productiva (Argentina),
% Minist\'{e}rio da Ci\^{e}ncia, Tecnologia e Inova\c{c}\~{a}o (Brazil),
% and Korea Astronomy and Space Science Institute (Republic of Korea).
% %
% Observations in the paper made use of the High-Resolution Imaging
% instrument Zorro at Gemini-South. Zorro was funded by the NASA
% Exoplanet Exploration Program and built at the NASA Ames Research
% Center by Steve B. Howell, Nic Scott, Elliott P. Horch, and Emmett
% Quigley.
% %
% This research has made use of the VizieR catalogue access tool, CDS,
% Strasbourg, France. The original description of the VizieR service was
% published in A\&AS 143, 23.
% %
% This work has made use of data from the European Space Agency (ESA)
% mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
% by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
% \url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
% for the DPAC has been provided by national institutions, in particular
% the institutions participating in the {\it Gaia} Multilateral
% Agreement.
%
% (Some of) The data presented herein were obtained at the W. M. Keck
% Observatory, which is operated as a scientific partnership among the
% California Institute of Technology, the University of California and
% the National Aeronautics and Space Administration. The Observatory was
% made possible by the generous financial support of the W. M. Keck
% Foundation.
% The authors wish to recognize and acknowledge the very significant
% cultural role and reverence that the summit of Maunakea has always had
% within the indigenous Hawaiian community.  We are most fortunate to
% have the opportunity to conduct observations from this mountain.
%
% \newline
%

\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  %\texttt{astroplan} \citep{astroplan2018},
  \texttt{astropy} \citep{astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  %\texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{cdips-pipeline} \citep{bhatti_cdips-pipeline_2019}
  \texttt{corner} \citep{corner_2016},
  %\texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{exoplanet} \citep{exoplanet:agol19}
  \texttt{exoplanet} \citep{exoplanet:exoplanet}, and its
  dependencies \citep{exoplanet:agol19, exoplanet:kipping13, exoplanet:luger18,
  	exoplanet:theano}.
  \texttt{IPython} \citep{perez_2007},
	\texttt{lightkurve} \citep{lightkurve_2018},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{MESA} \citep{paxton_modules_2011,paxton_modules_2013,paxton_modules_2015}
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{pyGAM} \citep{serven_pygam_2018_1476122},
  \texttt{PyMC3} \citep{salvatier_2016_PyMC3},
  %\texttt{radvel} \citep{fulton_radvel_2018},
  %\texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
  \texttt{tesscut} \citep{brasseur_astrocut_2019},
  \texttt{wotan} \citep{hippke_wotan_2019}.
}

\facilities{
 	{\it Astrometry}:
 	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}.
 	{\it Imaging}:
 	Keck:II~(NIRC2; \url{www2.keck.hawaii.edu/inst/nirc2})
 	%Gemini:South~(Zorro; \citealt{scott_nessi_2018}.
 	{\it Spectroscopy}:
 	Keck:I~(HIRES; \citealt{vogt_hires_1994}),
% 	Euler1.2m~(CORALIE),
% 	ESO:3.6m~(HARPS; \citealt{mayor_setting_2003}).
 	{\it Photometry}:
% 	CTIO:1.0m (Y4KCam),
% 	Danish 1.54m Telescope,
% 	El Sauce:0.356m,
% 	Elizabeth 1.0m at SAAO,
% 	Euler1.2m (EulerCam),
% 	Magellan:Baade (MagIC),
% 	Max Planck:2.2m	(GROND; \citealt{greiner_grond7-channel_2008})
% 	NTT,
% 	SOAR (SOI),
 	TESS \citep{ricker_transiting_2015}.
% 	TRAPPIST \citep{jehin_trappist_2011},
% 	VLT:Antu (FORS2).
}

@@

% \clearpage
\input{model_comparison_table.tex}
\input{model_posterior_table.tex}
% \clearpage
